\bf Calendar-based graphics for visualizing people's daily schedules

### 2019-05-15 reviews/response.Rmd

---
title: Response to reviewers
author: JCGS-18-293
output:
  MonashEBSTemplates::memo:
    includes:
      in_header: preamble2.tex
fontfamily: mathpazo
---

We thank the reviewers for their helpful feedback which has led to many improvements in the paper. The point by point description of changes are below: the reviewers' comments are in red and our response is in black.

In addition, we have expanded the paper by adding a subsection titled "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and a short section titled "3 Making calendar facets" which describes a new function `facet_calendar()`, that extends **ggplot2** intervals and complements the existing function `frame_calendar()`. It allows this algorithm to integrate more deeply with the **ggplot2** architecture.

@@

# Reviewer 1 comments {-}

> Figure 2 and Figure 10 show much the same information. You can see the Australian Open period in both. There is more detail in Figure 10 for the State Library.  The gaps in the Birrarung Marr data are easier to see in Figure 2.

We have explained the purpose of showing both displays in the revised paper, by adding a new subsection "2.4 Reasons to use calendar-based graphics" to explain why to use calendar plots, and what information is typically revealed by calendar plots but not by other plots. The calendar display doesn't replace traditional displays, but serves to complement them to further tease out interesting structures in the data.

@@

> Figure 11 is not described well.  What does the V-pattern mean?  How can you see the strong autocorrelation at weekends?  There are only dots.

We have changed the explanation of Fig 11 to make it clearer. In the revised paper this is now Fig 10. The V pattern is due to morning and evening commuter pattern on weekdays. We realised that the daily lag relationships would be read more easily if local scale was used, that is, each day is scaled relative to itself. This has been done, and the different autocorrelation pattern for week days and weekends is now clearer.

@@

> Do Figures 16-19 show much?  Drawing all four series as rows in facets one above the other would show the holidays better.  Comparisons of the households would be easier.  The six-weeks months confuse looking at holidays.  What are the one-day peaks in the middle of holidays?

Yes, we agree, they take up too much space. But the message of comparing multiple households is good to illustrate the benefits of the calendar plots, so we have reduced the plots to focus only on Jan-June 2018 and show all four households in one plot.

> There is a shiny app that presents some of the data.  It is not well described.  The plot at lower left is cumulative precipitation for each month, but it does not say so.  Why is cumulative precipitation by month interesting?  Months are of different lengths.  You can select single days in the time series.  Can you select more than one day?

We have added a text box underneath the plots in the shiny app. To describe the app in this paper would be too distracting for the reader. Cumulating monthly precipitation provides a more interesting look at rainfall, and it approximates how the NY Times displays weather. Only one day can be selected.

@@

> Fig.2: Can you match the minor grid lines with the start of a month, rather than the middle of a month? This will make it much easier to identify the approximate dates for some of the other spikes. No need to label these minor grid lines.

Done.

> Fig.2: You mention "small multiples" only in a figure caption. This concept is a central
part of your calendar graphics and should be summarized in more details (including references) in the main text. Also mention Unwin & Valero-Mora's "Ensemble Graphics", JCGS, 27(1), as a major concept that applies to both sets of related figures in your article.

We have explained the concept of faceting and also referred to small multiples, trellis charts, and ensemble graphics in a new paragraph in the  Introduction section.

@@

> Fig.4: Colors orange & purple are hard to distinguish in grayscale. Moreover, you should use different colors here as these 2 colors are already related to the 3 stations.

Done.

> Fig.4: Prior to reading the text on p8, I was really confused and even assumed there was a major bug in your R code. Having 2 days in early May and 1 day in early Oct, then 3 days missing, and then the remainder of the month is really confusing for someone who only looks at the figure without reading the text in detail. At least mention this layout anomaly in the figure caption.

Thanks. We have pointed out the layout anomalies for May and October in the Figure 4 (now 6) caption.

@@

> p13, possible subsection 2.2.7: How can you enter specific holidays? These are different from country to country, e.g., based on national holidays such as Thanksgiving, Labor Day, independence days, religious holidays, etc.

We used `tsibble::holiday_aus()` for generating Australian public holidays, prior to creating calendar displays. Whether to use holidays or not is a data-specific problem, and would depend on local analysts incorporating a holiday calculator into their country specific analyses. This is beyond the scope of this paper.

> Fig.8: "impossible to compare the size of peaks between days." - Correct; since this is based on ggplot, could you use colors from a sequential color scheme that maps the counts, e.g., from faint yellow (0) to dark red (for the overall max)? This might be another
useful argument for a future extension of your function.

We recommend against this idea. The better approach is to re-make the plot with global scaling to focus on counts. No need to make one plot show everything possible, but better to make multiple plots.

### 2019-05-15 main.Rmd
revised section 2.3

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  warning = FALSE, message = FALSE, echo = FALSE,
  fig.path = 'figure/', cache.path = 'cache/', fig.align = 'center',
  fig.show = 'hold', cache = FALSE, external = TRUE,
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
read_chunk('scripts/main.R')
```

@@

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \@ref(fig:facet) presents the idea of faceting calendar plots by sensors. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January. This is concealed in the conventional graphics.

@@

The `frame_calendar` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \@ref(fig:scatterplot) shows a lag scatterplot with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \@ref(fig:scatterplot) indicates two dominate patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days. At the higher counts on week days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs. This is consistent with the bimodality from Figure \@ref(fig:fs-2016), since positive correlations are perceived when approaching to the peak hour yet negative correlations when moving away from the peak hour.

>>>>

```{r initial, echo = FALSE, cache = FALSE, include = FALSE}
options("knitr.graphics.auto_pdf" = TRUE)
library(knitr)
opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE, comment = "#>",
  fig.path = 'figure/', fig.align = 'center', fig.show = 'hold', 
  cache = FALSE, cache.path = 'cache/', 
  out.width = ifelse(is_html_output(), "100%", "\\textwidth")
)
knitr::opts_knit$set(root.dir = here::here())
```

```{r external, include = FALSE}
read_chunk('scripts/main.R')
```

@@

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \@ref(fig:facet) presents the idea of faceting calendar plots by sensors. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September. The Australian Open, a major international tennis tournament, attracted much traffic constantly in the last two weeks of January.

@@

The `frame_calendar()` function is not constrained to line plots. The full range of plotting capabilities in **ggplot2** is essentially available. Figure \@ref(fig:scatterplot) shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the `x` argument and the current hourly count to the `y` argument. This figure is organized in the daily calendar layout. Figure \@ref(fig:scatterplot) indicates strong autocorrelation on weekends, and weak autocorrelation on work days. At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs. This is due to the morning and afternoon commuter patterns in Figure \@ref(fig:fs-2016): the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.

### 2019-05-15 main.tex
revised section 2.3

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of faceting calendar plots by sensors. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed, for example during the Australian Open, a major international tennis tournament, in the last two weeks of January. This is concealed in the conventional graphics.

@@

The \texttt{frame\_calendar} function is not constrained to line plots. The full range of plotting capabilities in \textbf{ggplot2} is essentially available. Figure \ref{fig:scatterplot} shows a lag scatterplot with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument. This figure is organized in the daily calendar layout. Figure \ref{fig:scatterplot} indicates two dominate patterns, strong autocorrelation on weekends, and weaker autocorrelation on work days. At the higher counts on week days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs. This is consistent with the bimodality from Figure \ref{fig:fs-2016}, since positive correlations are perceived when approaching to the peak hour yet negative correlations when moving away from the peak hour.

>>>>

To avoid the overlapping problem, the calendar layout can be embedded into a series of subplots for the different sensors. Figure \ref{fig:facet} presents the idea of faceting calendar plots by sensors. This allows comparing the overall structure between sensors, while emphasizing individual sensor variation. In particular, it can be immediately learned that Birrarung Marr was busy and packed over many weekends, but events took place on Friday evenings only in September. The Australian Open, a major international tennis tournament, attracted much traffic constantly in the last two weeks of January.

@@

The \texttt{frame\_calendar()} function is not constrained to line plots. The full range of plotting capabilities in \textbf{ggplot2} is essentially available. Figure \ref{fig:scatterplot} shows lag scatterplots with local scaling for each day at Flagstaff Station, where the lagged hourly count is assigned to the \texttt{x} argument and the current hourly count to the \texttt{y} argument. This figure is organized in the daily calendar layout. Figure \ref{fig:scatterplot} indicates strong autocorrelation on weekends, and weak autocorrelation on work days. At the higher counts on work days, the next hour sees possibly substantial increase or decrease in counts, as is depicted as the V-shape in the graphs. This is due to the morning and afternoon commuter patterns in Figure \ref{fig:fs-2016}: the adjacent hours are positively correlated when approaching to the peak hour but negatively correlated when moving away from the peak.
